---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
```{r}
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(lubridate))
suppressWarnings(library(zoo))
unzip("activity.zip")
data<-tbl_df(read.csv("activity.csv"))
```

##Make a histogram of the total number of steps taken each day
```{r}
datanaremove<-na.omit(data)
datanaremoveday<-summarise(group_by(datanaremove,date),sum(steps))
names(datanaremoveday)<-c("date","steps")
qplot(steps,data=datanaremoveday,fill=date)
```

##Calculate and report the **mean** and **median** total number of steps taken per day
```{r}
mean1<-mean(datanaremoveday$steps)
median1<-median(datanaremoveday$steps)
```
The mean and median of the total number of steps taken per day are `r mean1` and `r median1` respectively.

##Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
datanaremoveinterval<-summarise(group_by(datanaremove,interval),mean(steps))
names(datanaremoveinterval)<-c("interval","steps")
qplot(interval,steps,data=datanaremoveinterval,geom = "line")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxinterval<-datanaremoveinterval[which.max(datanaremoveinterval$steps),1]
```
The `r maxinterval` 5-minute interval contains the maximum number of steps.

## Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)
```{r}
sumna<-sum(is.na(data$steps))
```
The total number of missing values in the dataset is `r sumna`.

##filling the misssing values and creating a new data frame.
```{r}
data$steps<-na.aggregate(data$steps,data$interval)
```

##make a histogram of the total number of steps taken each day after imputing the NAs.
```{r}
dataday<-summarise(group_by(data,date),sum(steps))
names(dataday)<-c("date","steps")
qplot(steps,data=dataday,fill=date)
```

##Calculate and report the **mean** and **median** total number of steps taken per day after imputing the NAs.
```{r}
mean2<-mean(dataday$steps)
median2<-median(dataday$steps)
```
The new mean and median of the total number of steps taken per day are `r mean2` and `r median2` respectively. The new mean and median is slightly higher than the mean and median computed earlier (`r mean1` and `r median1` respectively). Hence, the estimates might be higher using the new dataset.

##Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
data$date<-as.character(data$date)
data$date<-ymd(data$date)
data<-mutate(data,weekday=ifelse(weekdays(data$date)=="Sunday"|weekdays(data$date)=="Saturday",0,1))
```

## Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
datainterval<-summarise(group_by(data,interval,weekday),mean(steps))
names(datainterval)<-c("interval","weekday","steps")
datainterval$weekday<-as.factor(datainterval$weekday)
levels(datainterval$weekday) <- c("weekends", "weekdays")
ggplot(data = datainterval,aes(interval,steps))+geom_line()+facet_grid(weekday~.)
```